---
title: "1. Determination of when to collect behavior"
author: "Justin Varholick"
date: '2022-06-14'
last modified: 
output: html_document
---

```{r setup}
library(tidyverse)
library(readxl)

library(ggplot2)
library(ggsci)
library(cowplot)
```

##Extract data from BORIS

1.  Open project file in BORIS

2.  Click the Observations Menu \> Export Events \> aggregated events

3.  Select the observations you wish to export, then all subjects and
    all behaviors

4.  Click YES for group events into one file

5.  Save as .csv file

One file for gg and one for jvselect all gg, and then only select the
necessary obs for jv

```{r import data and organize}
#import data
data_master <- read.csv("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/Ag_Boris_20211109_fromMaster_B1B2B3.csv")
not_reliability <- c("jv2", "gg2")

#organize
data_1 <- data_master %>% 
  separate(col=Observation.id, into=c("Cage", "Date", "Hour", "Coder"), sep="_") %>% 
  filter(!Coder %in% not_reliability) %>% 
  select(Cage, Date, Hour, Coder, Behavior, Start..s.)

head(data_1)
levels(as.factor(data_1$Coder))
data_1$Cage <- as.factor(data_1$Cage)
data_1$Date <- as.factor(data_1$Date)
data_1$Hour <- as.factor(data_1$Hour)
data_1$Coder <- as.factor(data_1$Coder)
levels(data_1$Cage)
levels(data_1$Date)
levels(data_1$Hour)
levels(data_1$Coder)
data_1 <- mutate(data_1, minute=round(Start..s./60))

#add sex ID
sex_id <- read_excel("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/AcomysDomRegen_Inventory.xlsx", sheet = "Sheet3")
sex_id$Cage <- as.factor(sex_id$Cage)

data_3 <- left_join(data_1, sex_id, by="Cage")

#add inactivity 
activity <- data_3[,c(1,2,3,8)]
activity <- unique(activity)
activity$Minute <- c(levels(data_1$minute))
```
##Plot activity across 24-hours
```{r activity plot}
batch1 <- c(957162, 922587, 941702, 927693)
batch1data <- data_3 %>% 
  filter(Cage %in% batch1) %>% 
  filter(Behavior=="Activity") %>% 
  group_by(Cage, Sex, Date, Hour) %>% 
  count()

zeromatrix_b1 <- data_3 %>% 
  filter(Cage %in% batch1) %>% 
  select(Cage, Sex, Date, Hour) %>% 
  distinct() %>% 
  mutate(n_prime=0)

batch1data <- left_join(zeromatrix_b1, batch1data, by=c("Cage", "Sex", "Date", "Hour"))
batch1data[is.na(batch1data)]<-0
batch1data <- batch1data[,c(1:4,6)]

batch1data <- mutate(batch1data, propact=(n/15)*100)
batch1data <- mutate(batch1data, species="Acomys")
batch1data$Hour <- factor(batch1data$Hour,levels=c("20","21","22","23","00","01","02","03","04","05","06","07","08","09","10","11","12","13","14","15","16","17","18","19"))

activity_plot <- ggplot(batch1data, aes(Hour, propact, group=species, color=Sex, shape=Sex)) +
  #stat_summary(fun=median, geom="line") +
  geom_pointrange(stat="summary",
                  fun.min=function(z) {quantile(z,0.25)},
                  fun.max=function(z) {quantile(z,0.75)},
                  fun=median)+
  geom_vline(xintercept="06", linetype="dashed", color="black")+
  ylab("Perc time active") + xlab("Hour of day")+
  scale_color_npg()+
  #scale_color_manual(values=c("#3498DB","#2ECC71")) +
  scale_x_discrete(labels=c("20","","","","00","","","","04","","","","08","","","","12","","","","16","","",""))+
  facet_grid(rows=vars(Sex))+
  guides(color="none", shape="none")
 
activity_plot
```
##Agonistic behavior during dark cycle
```{r agonistic night plot}
lightness <- c("06","07","08","09","10","11","12","13","14","15","16","17","18","19")
data_night <- data_3 %>% 
  filter(!Hour %in% lightness) %>% 
  filter(Cage %in% batch1)

not_ag <- c("Side Huddle", "Mounted Huddle", "Activity")
nightag <- data_night %>% 
  filter(!Behavior %in% not_ag) %>% 
  mutate(Behavior2="Agonistic") %>% 
  group_by(Cage, Sex, Date, Hour, Behavior2) %>% 
  count() %>% 
  mutate(prop=(n/15)*100)


zeromatrix <- data_night %>% 
  select(Cage, Sex, Date, Hour) %>% 
  distinct() %>% 
  mutate(Behavior2="Agonistic") %>% 
  mutate(n_prime=0) 

nightag <- left_join(zeromatrix, nightag, by=c("Cage", "Sex", "Date", "Hour", "Behavior2")) 
nightag[is.na(nightag)]<-0 
nightag <- nightag[,c(1:5,7:8)] 

nightag$Hour <- factor(nightag$Hour,levels=c("20","21","22","23","00","01","02","03","04","05"))

night_plot <- ggplot(nightag, aes(Hour, n, group=Sex, color=Sex, shape=Sex)) +
  geom_pointrange(stat="summary",
                  fun.min=function(z) {quantile(z,0.25)},
                  fun.max=function(z) {quantile(z,0.75)},
                  fun=median, 
                  position=position_dodge(width=0.5))+
  ylab("frequency ag bhv") + xlab("Hour of night")+
  scale_color_npg()+
  labs(color="Sex", shape="Sex")

night_plot
```

##Grouped plot and export
```{r grouped plot and export}
supp_Fig2 <- ggdraw() + 
  draw_plot(activity_plot, x=0, y=.5, width=1, height=.5) +
  draw_plot(night_plot, x=0, y=0, width=1, height=.5)+
  draw_plot_label(label=c("A","B"), size=12, x=c(0,0), y=c(1,.5))

ggsave("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/R generated figures/supp_Fig2.tiff", supp_Fig2, units="cm", width = 17, height = 14, device='tiff', dpi=300)

supp_Fig2
```

