---
title: "2. Differences ind-level social dom bhv"
author: "Justin Varholick"
date: '2022-06-14'
output: html_document
---

```{r setup}
library(tidyverse)
library(readxl)

library(ggplot2)
library(ggsci)
library(cowplot)
```
##Extract data from BORIS

1.  Open project file in BORIS

2.  Click the Observations Menu \> Export Events \> aggregated events

3.  Select the observations you wish to export, then all subjects and
    all behaviors

4.  Click YES for group events into one file

5.  Save as .csv file

One file for gg and one for jvselect all gg, and then only select the
necessary obs for jv

```{r data import}
#import data
data_master <- read.csv("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/Ag_Boris_20211109_fromMaster_B1B2B3.csv")
not_reliability <- c("jv2", "gg2")

#organize
data_1 <- data_master %>% 
  separate(col=Observation.id, into=c("Cage", "Date", "Hour", "Coder"), sep="_") %>% 
  filter(!Coder %in% not_reliability) %>% 
  select(Cage, Date, Hour, Coder, Behavior, Start..s.)

head(data_1)
levels(as.factor(data_1$Coder))
data_1$Cage <- as.factor(data_1$Cage)
data_1$Date <- as.factor(data_1$Date)
data_1$Hour <- as.factor(data_1$Hour)
data_1$Coder <- as.factor(data_1$Coder)
levels(data_1$Cage)
levels(data_1$Date)
levels(data_1$Hour)
levels(data_1$Coder)
data_1 <- mutate(data_1, minute=round(Start..s./60))

#add sex ID
sex_id <- read_excel("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/AcomysDomRegen_Inventory.xlsx", sheet = "sexes")
sex_id$Cage <- as.factor(sex_id$Cage)

data_3 <- left_join(data_1, sex_id, by="Cage")

#add inactivity 
activity <- data_3[,c(1,2,3,8)]
activity <- unique(activity)
activity$Minute <- c(levels(data_1$minute))

#limit to night
lightness <- c("06","07","08","09","10","11","12","13","14","15","16","17","18","19")
data_night <- data_3 %>% 
  filter(!Hour %in% lightness) 

not_ag <- c("Side Huddle", "Mounted Huddle", "Activity")
nightag <- data_night %>% 
  filter(!Behavior %in% not_ag) %>% 
  mutate(Behavior2="Agonistic") %>% 
  group_by(Cage, Sex, Date, Hour, Behavior2) %>% 
  count() %>% 
  mutate(prop=(n/15)*100)


zeromatrix <- data_night %>% 
  select(Cage, Sex, Date, Hour) %>% 
  distinct() %>% 
  mutate(Behavior2="Agonistic") %>% 
  mutate(n_prime=0) 

nightag <- left_join(zeromatrix, nightag, by=c("Cage", "Sex", "Date", "Hour", "Behavior2")) 
nightag[is.na(nightag)]<-0 
nightag <- nightag[,c(1:5,7:8)] 

nightag$Hour <- factor(nightag$Hour,levels=c("20","21","22","23","00","01","02","03","04","05"))


```
##Division of experimental groups
```{r filter exp groups}
sexdyads <- nightag

```

