---
title: "2. Social dominance status and steepness"
author: "Justin Varholick"
date: '2022-06-15'
output: html_document
---

```{r setup}

#plots
library(tidyverse)
library(ggplot2)
library(ggbeeswarm)
library(ggthemr) #devtools::install_github('Mikata-Project/ggthemr')
library(ggsci)
library(ggpubr)
library(cowplot)
library(readxl)

#dominance scripts
library(lubridate)
library(igraph)
library(steepness)

#stats
library(psych)
library(lme4)
library(lmerTest)
library(MuMIn)

ggthemr('flat')
```

```{r jalapic compete dominance scripts, modified by javarhol/author}
get_wl_dataframe <- function(df){
  library(dplyr)
  colnames(df)[1] <- "id1"
  colnames(df)[2] <- "id2"
  colnames(df)[3] <- "result"
  
  #df[,3] <- gsub("W", "1", df[,3])
  #df[,3] <- gsub("L", 2, df[,3])
  
  #df[,3] <- as.numeric(unlist(df[,3]))
  
  #turn into winner-loser dataframe
  df <- df[!is.na(df[,3]), ]  #get rid of NAs ties if exist
  a1 <- df %>%
    dplyr::filter(result==1) %>%
    dplyr::rename(winner=id1, loser=id2) %>%
    dplyr::mutate(result="W")
  
  a2 <- df %>%
    dplyr::filter(result==2) %>%
    dplyr::rename(winner=id2, loser=id1) %>%
    dplyr::mutate(result="W")
  
  a3 <- df %>%
    dplyr::filter(result==0) %>%
    dplyr::rename(winner=id1, loser=id2) %>%
    dplyr::mutate(result="T")
  df <-  rbind(a1,a2,a3)
  
  return(df)
}

get_wl_matrixold <- function(df){
    df <- get_wl_dataframe(df)
    df <- as.matrix(df[,1:2])
    g<-igraph::graph.edgelist(df)
    m1 <- igraph::get.adjacency(g, sparse=FALSE)
    m1 <- m1[order(rownames(m1)), order(colnames(m1))]
    return(m1)
}

ds = function(m, norm=FALSE, type="D"){
  M = as.matrix(m)
  diag(M)=0
  M1 = M + t(M)
  if(type=="D") { M2 = M/M1 - (((M/M1) - 0.5)/(M1 + 1)) }
  if(type=="P") { M2 = M/M1 }
  M2[is.na(M2)]=0
  ds = as.vector(rowSums(M2) + ( M2 %*% rowSums(M2)) - colSums(M2) - t(t(colSums(M2)) %*% M2))
  if(norm==TRUE){ ds = (ds + rshps(m)[[1]])/nrow(M)}
  names(ds)=colnames(m)
  return(ds)
}

rshps=function(m){
  M = as.matrix(m)
  diag(M)=0
  totalr = (nrow(M)*(nrow(M)-1))/2
  unk = unknowns(M)
  v=M[M==t(M)]
  ties=length(v[v>0])/2
  twoway= (length( M[M>0 & t(M)>0] )/2) - ties
  oneway = totalr-unk-ties-twoway
  return(list('total'=totalr, 'unknowns'=unk, 'ties'=ties, 'twoways'=twoway, 'oneways'=oneway))
}

unknowns<-function(m){
  m<-as.matrix(m)
  unk<-(length(m[m==0 & t(m)==0])-length(diag(m)))/2
  return(unk)
}

DScore <- function(x){
  matrix <- get_wl_matrixold(x)
  dscore <- as.data.frame(ds(matrix, norm=FALSE, type="D"))
}

getMinRank <- function(x){
  matrix <- get_wl_matrixold(x)
  Wins <- rowSums(matrix)
  TotalWins <- as.data.frame(Wins)
  TotalWins$ID <- rownames(TotalWins)
  TotalWins <- mutate(TotalWins, rank = rank(-Wins, ties.method="min"))
  TotalWins <- select(TotalWins, ID, Wins, rank)
  return(TotalWins)
}

getWins <- function(x){
  matrix <- get_wl_matrixold(x, ties="keep")
  Wins <- as.data.frame(matrix)
  Wins <- filter(Wins, Freq!=0)
  return(Wins)
}

DC <- function(x){
  matrix <- get_wl_matrixold(x)
  direccons<- as.data.frame(dc_compute(matrix))
}

dcmatrix=function(Matrix){
  Matrix=as.matrix(Matrix)
  diag(Matrix)=0
  DC=(abs(Matrix-t(Matrix)))/(Matrix+t(Matrix))
  return(DC)
}

DC_Matrix <- function(x){
  Matrix <- get_wl_matrix(x)
  dcmatrix(Matrix)
}

dc_compute=function(Matrix){
  Matrix=as.matrix(Matrix)
  diag(Matrix)=0
  N=sum(Matrix)/2
  DC=sum(abs(Matrix-t(Matrix)))/2/sum(Matrix)
  S=(Matrix+t(Matrix))/2
  K=(Matrix-t(Matrix))/2
  phi=sum(diag(t(K)%*%K))/sum(diag(t(Matrix)%*%Matrix))
  #phi is the skew-symmetrical index
  si=1-phi
  #si is the symmetrical index
  #result=list(DC=DC,S=S,K=K,phi=phi,si=si)
  #return(result)
  return(DC)
}

triangle <- function(x){
  m <- get_wl_matrixold(x, ties = "keep")
  mat <- get_di_matrix(as.matrix(m))
  #diag(mat)=0
  g = igraph::graph.adjacency(mat, mode="directed", diag=FALSE)
  tri=igraph::triad.census(g)
  return(which(tri == 1))
}

Steep <- function(x){
  matrix <- get_wl_matrixold(x)
  Steepdata <- as.data.frame(getStp(matrix, method = c("Dij","Pij")))
}

NormDS <- function(x){
  matrix <- get_wl_matrixold(x)
  normdsdata <- as.data.frame(getNormDS(matrix, method = c("Dij","Pij")))
}
```

##Extract data from BORIS
1 Open project file in BORIS
2 Click the Observations Menu > Export Events > aggregated events
3 Select the observations you wish to export, then all subjects and all behaviors
4 Click YES for group events into one file
5 Save as .csv file
6 edit in excel 
  1. split by "_" delimiter into 6 columns
  2. remove "MVI" from data
  3. Check all dates are actual dates (remove "2020")
  4. check all hours are actual hours (replace pt1, and shift 21 and 12 which extend to next column)
  5. check that all columns have the correct data, delete the coder initials data

```{r import data}
adata <- read.csv("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/dominanceevents_MASTER_adj.csv", stringsAsFactors=TRUE) 

#Fixing observation.id inconsistencies...
adata$Cage <- adata$Observation.id
adata$Subject <- gsub("Mouse ","", as.factor(adata$Subject))
adata$Date <- gsub("-Jan","01", adata$Date)
adata$Modifiers <- gsub("None", "3", adata$Modifiers)
adata$Date <- gsub("3001", "20200130", adata$Date)
adata$Date <- gsub("2901", "20200129", adata$Date)
adata$Date <- gsub("2801", "20200128", adata$Date)
adata$Date <- gsub("2301", "20200123", adata$Date)
adata$Date <- gsub("2201", "20200122", adata$Date)
adata$Date <- gsub("2101", "20200121", adata$Date)
adata$Date <- gsub("730", "20190730", adata$Date)
adata$Date <- gsub("731", "20190731", adata$Date)
adata$Date <- gsub("801", "20190801", adata$Date)
adata$Date <- gsub("806", "20190806", adata$Date)
adata$Date <- gsub("807", "20190807", adata$Date)
adata$Date <- gsub("808", "20190808", adata$Date)
adata$Date <- gsub("815", "20190815", adata$Date)
adata$Date <- gsub("816", "20190816", adata$Date)
adata$Date <- gsub("817", "20190817", adata$Date)
adata$Date <- gsub("2019020190731", "20190731", adata$Date)
adata$Date <- gsub("2019020190807", "20190807", adata$Date)
adata$Date <- gsub("916", "20200916", adata$Date)
adata$Date <- gsub("922", "20200922", adata$Date)
adata$Date <- gsub("923", "20200923", adata$Date)
adata$Date <- gsub("2020020200916", "20200916", adata$Date)
adata$Date <- gsub("2020020200922", "20200922", adata$Date)
adata$Date <- gsub("2020020200923", "20200923", adata$Date)

adata$Hour2 <- paste(adata$Hour, "00", sep=":")
adata$Date <- ymd(adata$Date)
adata$DateHour <- paste(adata$Date, adata$Hour2, sep=" ")
adata$DateHour <- as.POSIXct(adata$DateHour, tz="America/New_York")
adata$DateHour <- ymd_hms(adata$DateHour)

adata$Start..s. <- duration(adata$Start..s., "seconds")
adata$Stop..s. <- duration(adata$Stop..s., "seconds")

adata$Start <- (adata$DateHour + adata$Start..s.)
adata$Stop <- (adata$DateHour + adata$Stop..s.)

```


## Data description
There are 3 behavioral categories: Agonistic, Location, and Other. Other refers to inactive or unseen. For dominance ranks you only need the Agonistic behaviors. 

```{r pull agonistic behaviors out, organize for dominance rank}
ag <- adata %>% 
  filter(Behavioral.category == "Agonistic") %>% 
  filter(Behavior!="Fleeing") %>% 
  filter(Behavior!="Freeze") %>% 
  within(Opponent.1 <- paste(Cage, Subject, sep="-")) %>% 
  within(Opponent.2 <- paste(Cage, Modifiers, sep="-")) %>% 
  within(value <- paste("1")) %>% 
  select(Opponent.1, Opponent.2, value, Date, Behavior, Cage, Week)
ag_all <- adata %>% 
  filter(Behavioral.category!="Location") %>% 
  within(Opponent.1 <- paste(Cage, Subject, sep="-")) %>% 
  within(Opponent.2 <- paste(Cage, Modifiers, sep="-")) %>% 
  within(value <- paste("1")) %>% 
  select(Cage, Opponent.1, Opponent.2, value, Date, Week, Behavior, Duration..s.)
```

#### Induced flees, need to be flipped
- The induced flee behavior was coded so that if mouse 2 ran from mouse 1, then 2 was the subject and 1 was the modifier. In the case that we are observing "wins" (but really should be losses) then we need to flip Opponent.1 and Opponent.2
```{r}
ag2 <- ag %>% 
  filter(Behavior=="Induced Flee") %>% 
  mutate(Opponent.1f=Opponent.2) %>% 
  mutate(Opponent.2f=Opponent.1) %>% 
  select(Opponent.1f, Opponent.2f, value, Date, Behavior, Cage, Week)

ag_wo <- filter(ag, Behavior!="Induced Flee")

ag_adj <- ag2 %>% 
  mutate(Opponent.1=Opponent.1f) %>% 
  mutate(Opponent.2=Opponent.2f) %>% 
  select(Opponent.1, Opponent.2, value, Date, Behavior, Cage, Week) %>% 
  rbind(ag_wo)
```


```{r split data for runs}
Daylist <- split(ag_adj, ag_adj$Date)
Cagelist <- split(ag_adj, ag_adj$Cage)

ag_adj$CageWeek <- paste(ag_adj$Cage, ag_adj$Week, sep="-Week")
Weeklist <- split(ag_adj, ag_adj$CageWeek)

```

```{r moment of truth}
dscore_cage <- do.call(rbind, lapply(Cagelist, DScore))
dscore_cage
dc_cage <- do.call(rbind, lapply(Cagelist, DC))
dc_cage

dscore_week <- do.call(rbind, lapply(Weeklist, DScore))

steep_cage <- do.call(rbind, lapply(Cagelist, Steep))

####Wins
inventory <- read_xlsx("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/AcomysDomRegen_Inventory.xlsx", sheet="Sheet2")

wins <- ag_adj %>% 
  group_by(Opponent.1) %>% 
  count(value)
wins <- wins %>% 
  left_join(ag_adj, by="Opponent.1") %>% 
  mutate(wins=n) %>% 
  select(Cage, Opponent.1, wins) %>% 
  distinct()
inventory2 <- inventory %>% 
  filter(Week=="1") %>% 
  mutate(Opponent.1=AnimalID) %>% 
  mutate(Cage=CageID) %>% 
  select(Opponent.1, Cage)
wins_rank <- inventory2 %>% 
  left_join(wins, by="Opponent.1") %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  group_by(Cage.x) %>% 
  mutate(dom_rank=rank(-wins, ties.method="average")) %>% 
  select(Opponent.1, Cage.x, wins, dom_rank)

wins_rank$dom_rank <- gsub("1.5", "0", wins_rank$dom_rank)
wins_rank$dom_status <- as.factor(wins_rank$dom_rank)
wins_rank$dom_status <- gsub("0", "Unclear", wins_rank$dom_status)
wins_rank$dom_status <- gsub("1", "Dominant", wins_rank$dom_status)
wins_rank$dom_status <- gsub("2", "Subordinate", wins_rank$dom_status)
wins_rank$dom_status <- gsub("3", "Subordinate", wins_rank$dom_status)

wins_rank

wins_week <- ag_adj %>% 
  group_by(Opponent.1, Week) %>% 
  count(value) %>% 
  mutate(wins=n)
inventory3 <- inventory %>% 
  mutate(Opponent.1=AnimalID) %>% 
  mutate(Cage=CageID) %>% 
  select(Opponent.1, Cage, Week)
wins_week2 <- inventory3 %>% 
  full_join(wins_week, by=c("Opponent.1", "Week")) %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  group_by(Cage, Week) %>% 
  mutate(dom_rank=rank(-wins, ties.method="average")) %>% 
  select(Opponent.1, Cage, Week, wins, dom_rank)

wins_week2$dom_rank <- gsub("1.5", "0", wins_week2$dom_rank)
wins_week2$dom_status <- as.factor(wins_week2$dom_rank)
wins_week2$dom_status <- gsub("1", "Dominant", wins_week2$dom_status)
wins_week2$dom_status <- gsub("2.5", "Subordinate", wins_week2$dom_status)
wins_week2$dom_status <- gsub("2", "Subordinate", wins_week2$dom_status)
wins_week2$dom_status <- gsub("3", "Subordinate", wins_week2$dom_status)
wins_week2$dom_status <- gsub("0", "Unclear", wins_week2$dom_status)

wins_week2
```

```{r}
write.csv(dscore_cage, "/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/dscorebyanimal.csv")
write.csv(dscore_week, "/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/dscorebweek.csv")
write.csv(wins_rank, "/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/winsbyanimal.csv")
write.csv(wins_week2, "/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/winsbyweek.csv")
write.csv(dc_cage, "/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/dcbycage.csv")
write.csv(steep_cage, "/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/steepbycage.csv")
```