---
title: "5. Differences ind-level social dom bhv"
author: "Justin Varholick"
date: '2022-06-14'
output: html_document
---

```{r setup}
library(tidyverse)
library(readxl)

library(lubridate)
library(igraph)

library(ggplot2)
library(ggsci)
library(cowplot)
library(ggthemr)
library(stringr)

ggthemr('flat')

#custom functions
small_wilcox <- function(group1, group2, grouping_variable, prediction, data){
  library(dplyr)
  data_temp <- dplyr::mutate(data, ranks=rank({{ prediction }}))
  data_gp1 <- dplyr::filter(data_temp, {{ grouping_variable }}=={{ group1 }})
  data_gp2 <- dplyr::filter(data_temp, {{ grouping_variable }}=={{ group2 }})
  group1sum <- sum(data_gp1$ranks)
  group2sum <- sum(data_gp2$ranks)
  total_n <- dplyr::count(data, {{ grouping_variable }})
  
  return(c(group1sum, group2sum, total_n))
}

```
##Extract data from BORIS
1 Open project file in BORIS
2 Click the Observations Menu > Export Events > aggregated events
3 Select the observations you wish to export, then all subjects and all behaviors
4 Click YES for group events into one file
5 Save as .csv file
6 edit in excel 
  1. split by "_" delimiter into 6 columns
  2. remove "MVI" from data
  3. Check all dates are actual dates (remove "2020")
  4. check all hours are actual hours (replace pt1, and shift 21 and 12 which extend to next column)
  5. check that all columns have the correct data, delete the coder initials data

```{r import data}
adata <- read.csv("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/dominanceevents_MASTER_adj.csv", stringsAsFactors=TRUE)

#Fixing observation.id inconsistencies...
adata$Cage <- adata$Observation.id
adata$Subject <- gsub("Mouse ","", as.factor(adata$Subject))
adata$Date <- gsub("-Jan","01", adata$Date)
adata$Modifiers <- gsub("None", "3", adata$Modifiers)
adata$Date <- gsub("3001", "20200130", adata$Date)
adata$Date <- gsub("2901", "20200129", adata$Date)
adata$Date <- gsub("2801", "20200128", adata$Date)
adata$Date <- gsub("2301", "20200123", adata$Date)
adata$Date <- gsub("2201", "20200122", adata$Date)
adata$Date <- gsub("2101", "20200121", adata$Date)
adata$Date <- gsub("730", "20190730", adata$Date)
adata$Date <- gsub("731", "20190731", adata$Date)
adata$Date <- gsub("801", "20190801", adata$Date)
adata$Date <- gsub("806", "20190806", adata$Date)
adata$Date <- gsub("807", "20190807", adata$Date)
adata$Date <- gsub("808", "20190808", adata$Date)
adata$Date <- gsub("815", "20190815", adata$Date)
adata$Date <- gsub("816", "20190816", adata$Date)
adata$Date <- gsub("817", "20190817", adata$Date)
adata$Date <- gsub("2019020190731", "20190731", adata$Date)
adata$Date <- gsub("2019020190807", "20190807", adata$Date)
adata$Date <- gsub("916", "20200916", adata$Date)
adata$Date <- gsub("922", "20200922", adata$Date)
adata$Date <- gsub("923", "20200923", adata$Date)
adata$Date <- gsub("2020020200916", "20200916", adata$Date)
adata$Date <- gsub("2020020200922", "20200922", adata$Date)
adata$Date <- gsub("2020020200923", "20200923", adata$Date)

adata$Hour2 <- paste(adata$Hour, "00", sep=":")
adata$Date <- ymd(adata$Date)
adata$DateHour <- paste(adata$Date, adata$Hour2, sep=" ")
adata$DateHour <- as.POSIXct(adata$DateHour, tz="America/New_York")
adata$DateHour <- ymd_hms(adata$DateHour)

adata$Start..s. <- duration(adata$Start..s., "seconds")
adata$Stop..s. <- duration(adata$Stop..s., "seconds")

adata$Start <- (adata$DateHour + adata$Start..s.)
adata$Stop <- (adata$DateHour + adata$Stop..s.)

ag <- adata %>% 
  within(Opponent.1 <- paste(Cage, Subject, sep="-")) %>% 
  within(value <- paste("1")) %>% 
  select(Opponent.1, value, Date, Behavior, Behavioral.category, Modifiers, Cage, Week, Start, Stop)

locations <- ag %>% 
  filter(Behavioral.category=="Location") %>% 
  filter(Behavior!="Shelter") %>% 
  filter(Behavior!="Food Hopper") %>% 
  filter(Behavior!="Cup") %>% 
  unite("Behavior2", Behavior, Modifiers, sep="-", remove=T) %>% 
  rename("Behavior"=Behavior2)

ag <- ag %>% 
  filter(Behavioral.category!="Location") %>% 
  select(Opponent.1, value, Date, Behavior, Behavioral.category, Cage, Week, Start, Stop)
ag <- rbind(ag, locations)

#add independent variables
sex_id <- read_excel("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/AcomysDomRegen_Inventory.xlsx", sheet = "Sheet3")
sex_id$Cage <- as.factor(sex_id$Cage)
ag$Cage <- as.factor(ag$Cage)

ag <- left_join(ag, sex_id, by="Cage")

ind_ds <- read.csv("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/dscorebyanimal_edited.csv")
ind_ds$Cage <- as.factor(ind_ds$Cage)
ind_ds <- rename(ind_ds, "Opponent.1"=AnimalID)

ag <- left_join(ag, ind_ds, by=c("Opponent.1", "Cage"))
```
##Do dominants engage in more frequency of chasing than other ag off bhvs?
```{r}

other_ag <- c("Attacking", "Mounting", "Stealing")

dom_data1 <- ag %>% 
  filter(dom_status_ds=="Dominant") %>% 
  filter(Behavior %in% other_ag) %>% 
  select(Opponent.1, Date, Start, Stop, Cage, Sex, GrpSize, Age, DumCage, dom_status_ds, value) %>% 
  within(Behavior <- paste("Other offensive \nBehaviors"))

dom_data2 <- ag %>% 
  filter(dom_status_ds=="Dominant") %>% 
  filter(Behavior=="Chasing") %>% 
  select(Opponent.1, Date, Start, Stop, Cage, Sex, GrpSize, Age, DumCage, dom_status_ds, value, Behavior)

dom_data <- rbind(dom_data1, dom_data2)

dom_data <- dom_data %>% 
  group_by(Opponent.1, Cage, Sex, GrpSize, Age, DumCage, Behavior) %>% 
  count(value)

zeromatrix <- ag %>% 
  filter(dom_status_ds=="Dominant") %>% 
  select(Opponent.1, Cage, Sex, GrpSize, Age, DumCage) %>% 
  unique() %>% 
  mutate(Chasing=0) %>% 
  mutate(`Other offensive \nBehaviors`=0) %>% 
  gather("Behavior", "n_prime", 7:8)

dom_data <- left_join(zeromatrix, dom_data, by=c("Opponent.1", "Cage", "Sex", "GrpSize", "Age", "DumCage", "Behavior"))
dom_data[is.na(dom_data)]<-0
dom_data <- dom_data[,c(1:7,10)]
dom_datax <- spread(dom_data, Behavior, n)
dom_datax <- mutate(dom_datax, test=abs(Chasing-`Other offensive \nBehaviors`))
dom_datax <- mutate(dom_datax, rank=rank(test))

sum(dom_datax$rank)-2

#43, 9, p=0.0059
```
```{r plot}
chase_plot <- ggplot(dom_data, aes(Behavior, n, group=as.factor(DumCage), color=Sex, fill=Sex)) +
  geom_point(size=3) + geom_line() +
  scale_color_npg() + scale_fill_npg()+
  ylab("total frequency") + xlab("")
chase_plot
```
#Is chasing higher in males than females?
```{r}
dom_chase <- filter(dom_data, Behavior=="Chasing")
small_wilcox("M", "F", Sex, n, dom_chase)
#W=16 (3, 9) p=0.300
```
#Higher in triads?
```{r}
group_chase <- filter(dom_chase, Sex=="M")
small_wilcox("2", "3", GrpSize, n, group_chase)
#W=8(3,3) p = 0.200
```


##Do subordinates engage in higher levels of fleeing than induced flee (displacement)?
```{r}
freeze <- filter(ag, Behavior=="Freeze") #(only 4 instances)

def <- c("Fleeing", "Induced Flee")

sub_data <- ag %>% 
  filter(dom_status_ds=="Subordinate") %>% 
  filter(Behavior %in% def) %>% 
  group_by(Opponent.1, Cage, Sex, GrpSize, Age, DumCage, Behavior) %>% 
  count(value)

zeromatrix <- ag %>% 
  filter(dom_status_ds=="Subordinate") %>% 
  select(Opponent.1, Cage, Sex, GrpSize, Age, DumCage) %>% 
  unique() %>% 
  mutate(Fleeing=0) %>% 
  mutate(`Induced Flee`=0) %>% 
  gather("Behavior", "n_prime", 7:8)

sub_data <- left_join(zeromatrix, sub_data, by=c("Opponent.1", "Cage", "Sex", "GrpSize", "Age", "DumCage", "Behavior"))
sub_data[is.na(sub_data)]<-0
sub_data <- sub_data[,c(1:7,10)]
sub_datax <- spread(sub_data, Behavior, n)
sub_datax <- mutate(sub_datax, test=abs(Fleeing-`Induced Flee`))
sub_datax <- mutate(sub_datax, rank=rank(test))

#sum of t- = 5
sum(sub_datax$rank)-5

#40, 9, 0.0195
```

```{r}
def <- c("Induced Flee", "Freeze")

sub_data <- ag %>% 
  filter(dom_status_ds=="Subordinate") %>% 
  filter(Behavior %in% def) %>% 
  group_by(Opponent.1, Cage, Sex, DumCage, Behavior) %>% 
  count(value)

zeromatrix <- ag %>% 
  filter(dom_status_ds=="Subordinate") %>% 
  select(Opponent.1, Cage, Sex, DumCage) %>% 
  unique() %>% 
  mutate(`Induced Flee`=0) %>% 
  mutate(Freeze=0) %>% 
  gather("Behavior", "n_prime", 5:6)

sub_data <- left_join(zeromatrix, sub_data, by=c("Opponent.1", "Cage", "Sex", "DumCage", "Behavior"))
sub_data[is.na(sub_data)]<-0
sub_data <- sub_data[,c(1:5,8)]
sub_datax <- spread(sub_data, Behavior, n)
sub_datax <- mutate(sub_datax, test=abs(`Induced Flee`-Freeze))
sub_datax <- mutate(sub_datax, rank=rank(test))

#no t-
sum(sub_datax$rank)

#45, 9, 0.0020
```
```{r}
def <- c("Fleeing", "Induced Flee", "Freeze")

sub_data <- ag %>% 
  filter(dom_status_ds=="Subordinate") %>% 
  filter(Behavior %in% def) %>% 
  group_by(Opponent.1, Cage, Sex, DumCage, Behavior) %>% 
  count(value)

zeromatrix <- ag %>% 
  filter(dom_status_ds=="Subordinate") %>% 
  select(Opponent.1, Cage, Sex, DumCage) %>% 
  unique() %>% 
  mutate(Fleeing=0) %>% 
  mutate(`Induced Flee`=0) %>% 
  mutate(Freeze=0) %>% 
  gather("Behavior", "n_prime", 5:7)

sub_data <- left_join(zeromatrix, sub_data, by=c("Opponent.1", "Cage", "Sex", "DumCage", "Behavior"))
sub_data[is.na(sub_data)]<-0
sub_data <- sub_data[,c(1:5,8)]

sub_data <- mutate(sub_data, Behavior=ordered(Behavior, levels=c("Fleeing", "Induced Flee", "Freeze")))
  
def_plot <- ggplot(sub_data, aes(Behavior, n, fill=Sex)) +
  geom_boxplot() +
  scale_color_npg() + scale_fill_npg()+
  ylab("total frequency") + xlab("")
def_plot
```


#Flee and induced flee between sexes
```{r}
sub_flee <- filter(sub_data, Behavior=="Induced Flee")
small_wilcox("M", "F", Sex, n, sub_flee)
#W(3,9)=15 p=.2409
```
```{r}
grp_flee <- filter(sub_flee, Sex=="M")
small_wilcox("2", "3", GrpSize, n, grp_flee)
```


##Total offensive vs. defensive behavior
```{r}
off <- c("Chasing", "Attacking", "Mounting", "Stealing")
def <- c("Fleeing", "Induced Flee", "Freeze")

dom_data <- ag %>% 
  filter(dom_status_ds=="Dominant") %>% 
  filter(Behavior %in% off) %>% 
  group_by(Opponent.1, Cage, Sex, DumCage) %>% 
  count(value) %>% 
  rename(total_off=n) 

sub_data <- ag %>% 
  filter(dom_status_ds=="Subordinate") %>% 
  filter(Behavior %in% def) %>% 
  group_by(Opponent.1, Cage, Sex, DumCage) %>% 
  count(value) %>% 
  rename(total_def=n)

#zeromatrix unnecessary or else they wouldn't be ranked.

domsub_data <- left_join(dom_data, sub_data, by=c("Cage", "Sex", "DumCage"))

domsub_data <- mutate(domsub_data, test=abs(total_def-total_off))
domsub_data <- mutate(domsub_data, rank=rank(test))

#t- = 1
#44, 9, p=0.0039
```

```{r}
domsub_datax <- domsub_data %>% 
  select(Cage, Sex, DumCage, total_off, total_def) %>% 
  gather(Behavior, n, 4:5)

domsub_plot <- ggplot(domsub_datax, aes(Behavior, n, group=as.factor(DumCage), color=Sex, fill=Sex)) +
  geom_point(size=3) + geom_line() +
  scale_color_npg() + scale_fill_npg()+
  scale_x_discrete(labels=c("Total Defensive \nfor Subordinates", "Total Offensive \nfor Dominants"))+
  ylab("total frequency") + xlab("")
domsub_plot
```
```{r grouped plot}
fig3 <- plot_grid(
  chase_plot, def_plot, domsub_plot,
  labels = "AUTO", ncol = 1
)

ggsave("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/R generated figures/Fig3_inddom_plots.tiff", fig3, units="cm", width = 7.5, height = 15, device='tiff', dpi=300)

  fig3
```


