---
title: "9. qPCR calculations"
author: "varholick"
date: "1/3/2022"
modified: "8/29/2022"
output: html_document

#Pfaffl Method
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readxl)
```

```{r import_data}
raw_data <- read_excel("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/qPCR_data_dominance1_rev.xlsx", sheet="DupsRemoved")
```
```{r double_check duplicates}
raw_data2 <- raw_data %>% 
  group_by(target, sample) %>% 
  mutate(num_dups=n(),
         dup_id=row_number()) %>%
  ungroup() %>% 
  mutate(isduplicated=dup_id > 1)
```
##### _check the number of duplicates, only calibrator should have more than 3_

### Log transform and calculate average Ct values
```{r tidy data}
tidy_data <- raw_data %>% 
  filter(sample!="NTC", sample!="") %>% 
  mutate(cq=as.numeric(cq)) %>% 
  mutate(logcq= log(cq)) %>% 
  group_by(plate, sample, target) %>% 
  summarise(mean_Ct = mean(logcq))
```

### Adjust Ct values for calibrator per plate (Calibrator - Ct)
```{r plate calibration}
test_data <- tidy_data %>% 
  filter(sample!="Calibrator")
calibrators <- tidy_data %>%
  filter(sample=="Calibrator")

cal_tidy <- left_join(test_data, calibrators, by=c("plate"))
cal_tidy <- mutate(cal_tidy, calb_Ct = mean_Ct.y - mean_Ct.x)
tidy_data_calCt <- cal_tidy %>% 
  data.frame() %>% 
  mutate(target=target.x) %>% 
  select(sample.x, target, calb_Ct)

head(tidy_data_calCt)
```

### Adjust Ct values for primer efficiency (removed on 04.11.2023)
Efficiences were determined on March 4th 2021 (Cyp11b1 on April 1st 2022)
With only three concentrations for calculation (1, .1, .01), extended concentrations in data folder
```{r primer efficiency}

#primer_effis <- read_xlsx("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/qPCR_primer_effis_converted.xlsx")

#tidy_data_calCt_eff <- left_join(tidy_data_calCt, primer_effis, by="target")

#tidy_data_calCt_eff <- tidy_data_calCt_eff %>%
  #mutate(eff_Ct = conv_effi^(calb_Ct))

#head(tidy_data_calCt_eff)
```

### Calculate DeltaCt (target - ActinB)
```{r}
test_data <- tidy_data_calCt %>% 
  filter(target!="ActinB")
ActinB <- tidy_data_calCt %>%
  filter(target=="ActinB")

HK_tidy <- left_join(test_data, ActinB, by=c("sample.x"))
HK_tidy <- mutate(HK_tidy, deltact = calb_Ct.x - calb_Ct.y)

tidy_data_deltact <- HK_tidy %>% 
  data.frame() %>% 
  select(sample.x, target.x, deltact)
head(tidy_data_deltact)
```

### Calculate delta delta Ct for each Gene of Interest 
- separate each gene of interest
- Divide delta Ct by the highest delta Ct value for each gene of interest
- 2^(-delta delta ct)
- merge each gene of interest to one dataframe
```{r}
#separate genes of interest
Cyp11a1 <- filter(tidy_data_deltact, target.x=="Cyp11a1")
Cyp11b1 <- filter(tidy_data_deltact, target.x=="Cyp11b1")
Nr5a1 <- filter(tidy_data_deltact, target.x=="Nr5a1")
Nrb01 <- filter(tidy_data_deltact, target.x=="Nrb01")
StAR <- filter(tidy_data_deltact, target.x=="StAR")

#highest delta ct for each gene of interest
Cyp11a1 <- mutate(Cyp11a1, highestdelta=max(deltact))
Cyp11b1 <- mutate(Cyp11b1, highestdelta=max(deltact))
Nr5a1 <- mutate(Nr5a1, highestdelta=max(deltact))
Nrb01 <- mutate(Nrb01, highestdelta=max(deltact))
StAR <- mutate(StAR, highestdelta=max(deltact))

#delta delta ct
Cyp11a1 <- mutate(Cyp11a1, deltadeltact=(highestdelta-deltact))
Cyp11b1 <- mutate(Cyp11b1, deltadeltact=(highestdelta-deltact))
Nr5a1 <- mutate(Nr5a1, deltadeltact=(highestdelta-deltact))
Nrb01 <- mutate(Nrb01, deltadeltact=(highestdelta-deltact))
StAR <- mutate(StAR, deltadeltact=(highestdelta-deltact))

#fold.change
Cyp11a1 <- mutate(Cyp11a1, rel_conc=(2^-(deltadeltact)))
Cyp11b1 <- mutate(Cyp11b1, rel_conc=(2^-(deltadeltact)))
Nr5a1 <- mutate(Nr5a1, rel_conc=(2^-(deltadeltact)))
Nrb01 <- mutate(Nrb01, rel_conc=(2^-(deltadeltact)))
StAR <- mutate(StAR, rel_conc=(2^-(deltadeltact)))

#merged data
rel_conc_data = rbind(Cyp11a1, Cyp11b1, Nr5a1, Nrb01, StAR)
```

### Add independent variables
```{r import inventory data}
inv_data <- read_excel("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/AcomysDomRegen_Inventory.xlsx")
```
```{r merge inv with deltaCt values}
tidy_inv_data <- inv_data[,c(1,2,5,6,7,8)]
rel_conc_data <- rename(rel_conc_data, BlindID = sample.x)
rel_conc_data2 <- left_join(rel_conc_data, tidy_inv_data, by="BlindID")
head(rel_conc_data2)
```

```{r}
write.csv(rel_conc_data2, "/Users/justinvarholick/Documents/GitHub/AcomysDominance_2022_Data_Results/Data/qpcr_combined_041123.csv")
          
```